---
title: "LLM-autocorprophagy"
subtitle: "Predictions, retrodictions and prognoses of LLMs’ impact on written language."
author: "Aleksander Dietrichson, PhD"
format: 
  revealjs:
    self-contained: true
  beamer: default
output-file: index
editor: visual
---


```{r}
#| include: false
library(tidytext)
library(tidyverse)
source(here::here("R","00-db-helpers.R"))
source(here::here("R/unique_word_count.R"))
con <- dbConnect(SQLite(), DB_PATH)
my_data <- read_cc_samples(con)
```

## Disclaimer & Reverse Disclaimer

-   *Claude code* was used to generate the source code for data collection and for running the experiment as reported.

-   No AI tools were used in the creation of this presentation. R and quarto code for data analysis written exclusively written by the author[^1].

-   Source-code, AI prompts and data available at: <https://github.com/dietrichson/ILA-2026>

[^1]: With the exception of spell-checker.

## How LLMs Work

```{mermaid}
flowchart TD
    A[Scrape Internet] --> B[Compression/Vectorization]
    B --> E[Model Training ]
    E --> F[Model Weights]
    F --prompt--> G[Output]
```

## The Problem

When most training data for LLMs is the output of other LLMs a loop is created

```{mermaid}
flowchart TD
    A[Scrape Internet] --> B[Compression/Vectorization]
    B --> E[Model Training ]
    E --> F[Model Weights]
    F --prompt: create web content--> G[Output]
    G -- becomes training data --> A
```


## Experiment: Data

-   Five articles, 1600-2200 words each.

-   One from the *New Yorker* four from the *New York Times*.

-   Chosen for length, from the latest edition of each publication (last week of april 2026)

## Experiment: Method

### Step 1: Summarization task

This is the text of an published article. Your task is to summarize this. Aim for 500 words.
The output filename should have the same name as the input-file with "_summary" interjected.

### Step 2: Expansion task

This is a short article, we would like to make it longer. Please add verbiage so that the article is around 1600 words long.
The output filename should be the same as the input-file with "_expanded" instead of "_summary"

## Experiment: Aparatus

-  Claude code was used, with Opus 4.7 


## Results

```{r}
#| echo: false
#| label: loss_table
#| tbl-cap: "Unique words (types) per condition"
publication_names <- c("New Yorker", "New York Times 1", "New York Times 2", "New York Times 3", "New York Times 4")
my_filenames_origininal <- c(
  here::here("data/articles/new_yorker_1.txt"),
  here::here("data/articles/nyt_1.txt"),
  here::here("data/articles/nyt_2.txt"),
  here::here("data/articles/nyt_3.txt"),
  here::here("data/articles/nyt_4.txt")
)
my_filenames_summary <- c(
  here::here("data/articles/new_yorker_1_summary.txt"),
  here::here("data/articles/nyt_1_summary.txt"),
  here::here("data/articles/nyt_2_summary.txt"),
  here::here("data/articles/nyt_3_summary.txt"),
  here::here("data/articles/nyt_4_summary.txt")
)
my_filenames_expanded <- c(
  here::here("data/articles/new_yorker_1_expanded.txt"),
  here::here("data/articles/nyt_1_expanded.txt"),
  here::here("data/articles/nyt_2_expanded.txt"),
  here::here("data/articles/nyt_3_expanded.txt"),
  here::here("data/articles/nyt_4_expanded.txt")
)

my_data_experiment <- data.frame(
  Publication = publication_names,
  original = purrr::map_dbl(my_filenames_origininal, unique_word_count),
  summary = purrr::map_dbl(my_filenames_summary, unique_word_count),
  expanded = purrr::map_dbl(my_filenames_expanded, unique_word_count)
)

loss <- my_data_experiment$original - my_data_experiment$expanded 
loss <- loss/my_data_experiment$original


my_data_experiment$Loss <- scales::percent(loss)
my_data_experiment |> 
  select(-summary) |> 
  knitr::kable(col.names = c("Publication", "Original", "Processed", "Loss"))

```

Average loss is `r scales::percent(mean(loss))`

## Results

```{r}
#| label: loss_fig
#| fig-cap: "Unique words (types) per condition"
my_data_experiment |> 
  select(Publication, original, expanded) |> 
  pivot_longer(cols = c(original, expanded)) |> 
  ggplot(aes(name, value, color = Publication, group = Publication)) +
  geom_point() +
  geom_line() +
  scale_x_discrete(limits = c("original", "expanded"))
```

## Empirical Research

-  Random sample of 15 thousand webpages
-  Retrieved from https://commoncrawl.org/
-  English language
-  1000 from each of the years 2012 through 2026
-  Calculated Type/Token ratio for each document

```{r}
my_data_tokenized <- my_data |>
  unnest_tokens(word, text)
  
# Get unique words per year
# my_data_tokenized |> 
#   group_by(year) |> 
#   summarize(
#     n_words = n_distinct(word)
#   )


TTRS <- my_data_tokenized |>
  group_by(id, year) |>
  summarize(
    n_tokens = n(),
    n_types  = n_distinct(word),
    ttr      = n_types / n_tokens,
    avg_word_length = mean(nchar(word), na.rm = TRUE)
  )


```

## Type Token Ratio

$$ 
TTR = {\text{number of unique words} \over {\text{total number of words}}}
$$

A measure of *Vocabulary diversity* in text

## Results

```{r}
#| label: "median_TRR"
#| fig-cap: "Median TRR over Year"

TTRS |> 
  group_by(year) |> 
  summarize(
    Median = median(ttr)
  ) |> 
  ggplot(aes(year,Median))+
  geom_point()+
  geom_line()+
  geom_smooth(method = "lm") -> my_plot

my_plot
```

## Results

```{r}
#| label: "mean_word_length"
#| fig-cap: "Mean Length of Words Over Year"

TTRS |> 
  group_by(year) |> 
  summarize(
    Mean = mean(avg_word_length)
  ) |> 
  ggplot(aes(year,Mean))+
   geom_point()+
   geom_line()+
  geom_smooth(method = "lm", se = FALSE, fullrange = TRUE) -> my_plot2
my_plot2
```


## Results

```{r}
#| include: false
TTRS |> 
  group_by(year) |> 
  summarize(
    Median = median(ttr)
  ) |> 
  mutate(change = lag(Median)-Median) -> tmp
TTRS |> 
  group_by(year) |> 
  summarize(
    Mean = mean(avg_word_length)
  ) |> mutate(change = Mean - lag(Mean)) -> tmp2
```

-  Type/Token ratio drops by `r mean(tmp$change, na.rm = TRUE)` (~ .8%) anually.
-  Word-length increases by `r tmp2$change |> mean(na.rm=TRUE)` character per word anually.

## Projections

```{r}
#| label: "mean_types"
#| fig-cap: "Mean Types (Raw Scores)"
TTRS |> 
  group_by(year) |> 
  summarize(
    Types = mean(n_types)
  ) |> 
  ggplot(aes(year,Types))+
  # geom_point()+
  # geom_line()+
  geom_smooth(method = "lm", se=FALSE, fullrange=TRUE) -> my_plot
my_plot + 
  scale_x_continuous(limits = c(1900, 2200))
```

## Conclusions

-  Linguistic diversity is under preassure. 
-  "AI slop" seems omnipresent and still feature recognizable characteristics



