---
title: "LLM-autocorprophagy"
subtitle: "Predictions, retrodictions and prognoses of LLMs’ impact on written language."
author: "Aleksander Dietrichson, PhD"
format: revealjs
editor: visual
---

## 

```{r}
#| include: false
library(tidytext)
library(tidyverse)
source(here::here("R","00-db-helpers.R"))
con <- dbConnect(SQLite(), DB_PATH)
my_data <- read_cc_samples(con)
```

## Disclaimer & Reverse Disclaimer

-   *Claude code* was used to generate the source code for data collection and for running the experiment as reported.

-   No AI tools were used in the creation of this presentation. R and quarto code for data analysis written exclusively written by the author[^1].

-   Source-code, AI prompts and data available at: <https://github.com/dietrichson/ILA-2026>

[^1]: With the exception of spell-checker.

## How LLMs Work

## Experiment 

```{r}

my_filename <- here::here("data/articles/new_yorker_1.txt")
my_text <- readLines(my_filename)
text_df <- tibble(line = 1:length(my_text), text = my_text)
text_df |> 
  unnest_tokens(word,text) |> 
  nrow()

text_df |> 
  unnest_tokens(word, text) |> 
  count(word, sort = TRUE) |> 
  nrow()

my_filename <- here::here("data/articles/new_yorker_1_expanded.txt")
my_text <- readLines(my_filename)
text_df <- tibble(line = 1:length(my_text), text = my_text)
text_df |> 
  unnest_tokens(word, text) |> 
  count(word, sort = TRUE) |> 
  nrow()

```

## Empirical Exploration
